The Weakly Nonlinear Magnetorotational Instability in a Global, Cylindrical Taylor-Couette Flow

### 2016-10-04 paper/widegap_mri_paper.tex
two ways to compute nablasq
minor changes.

%\documentclass[onecolumn]{emulateapj}
\documentclass{emulateapj}
\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}

@@

\usepackage[colorlinks,urlcolor=blue,citecolor=blue,linkcolor=blue]{hyperref}
\usepackage{gensymb}
\usepackage{amsmath}
\usepackage{color}
%\usepackage{appendix}
\usepackage[titletoc,title]{appendix}
\usepackage{cleveref}

@@

\begin{document}
%\title{Weakly Nonlinear Analysis of the Standard and Helical Magnetorotational Instabilities in a Taylor Couette Flow}
\title{The Weakly Nonlinear Magnetorotational Instability in a Global, Cylindrical Taylor-Couette Flow} % maybe this is punchier?
\author{Clark, S.E.\altaffilmark{1}}
\author{Oishi, J.S. \altaffilmark{2,}\altaffilmark{3}}
\altaffiltext{1}{Department of Astronomy, Columbia University, New York, NY} 
\altaffiltext{2}{Department of Astrophysics, American Museum of Natural History, New York, NY}
\altaffiltext{3}{Department of Physics and Astronomy, Bates College}

@@

%\begin{widetext}
\begin{table*}[ht]
\normalsize
\caption{Fiducial parameters for MRI runs} \label{table:parameters}
\centering
\begin{tabular}{rlrrrrrrr}
  \hline
 & $\xi$ & $\Pm$ & $\beta$ & $\Omega_2/\Omega_1$ & $R_1/R_2$ & radial magnetic b.c. \\ 
  \hline\hline
Standard MRI & 0 & 1.6E-6 & 41.2 & 0.121 & 0.33 & conducting \\ 
Helical MRI & 4 & 1E-6 & 1.7E-2 & 0.27 & 0.5 & insulating\\ 
   \hline
\end{tabular}
\end{table*} 
%\end{widetext}

@@

%Because of the MRI's complexity, analytical treatments of the MRI use some combination of convenient approximations, such as ideal MHD, linearized equations, expedient boundary conditions, and simplified geometries. In particular, much of the analytical work on MRI saturation has focused on local approximations, often representing a section of a disk by solving the MHD equations in an isolated region subject to shear periodic boundary conditions (the ``shearing box"). The shearing box is an inexpensive computational framework, so is used by many numerical MRI studies as well. However, while the MRI is a local instability, certain properties of the local problem are not generic to the global problem. In the shearing box, linear evolution is dominated by a class of MRI mode known as channel modes.  %Channel modes, a class of MRI mode that dominates MRI evolution and saturation in the local approximation, are artificially over-prominent in the shearing box (e.g. \citei{Latter:2015}). 
%These are linear modes which also happen to be exact solutions to the \textit{nonlinear} local MRI equations. Runaway growth is avoided in this paradigm by the instability of the channel modes themselves, which are destroyed by parasitic (secondary) instabilities (\citei{Goodman:1994ul}, \citei{Pessah:2010ic}). The growth of parasitic modes provides a saturation avenue for channel mode-dominated flows, yet this is unlikely to be the dominant saturation mechanism in laboratory experiments or astrophysical disks, as channel modes are artificially over-prominent in the shearing box (e.g. \citei{Latter:2015}).

@@

\beq\label{momentum}
%\begin{split}
\partial_t \mathbf{u} + \mathbf{u} \cdot \nabla \mathbf{u} = -\frac{1}{\rho}\nabla P - \nabla\Phi + \frac{1}{c\rho} \left(\mathbf{J}\times\mathbf{B}\right) + \nu\nabla^2 \mathbf{u} 
%\end{split}
\eeq

@@

%and define the magnetic vector potential $A$ analogously. 
These definitions automatically satisfy Equations \ref{eq:incompressibility} and \ref{eq:solenoid} for axisymmetric disturbances. We note that in the linearized equations, streamfunctions of the form $u_x = \partial_z \Psi$, $u_z = -(\partial_r + \frac{1}{r}) \Psi$, and the corresponding definitions of the magnetic vector potential, are convenient choices, but we define Equations \ref{eq:stokes} and \ref{eq:fluxfunc} for this nonlinear investigation because of the incommutability of $\partial_r$ and $\partial_r + \frac{1}{r}$.

@@

\begin{align}
%\begin{split}
& \frac{1}{r}\partial_t (\nabla^2 \Psi - \frac{2}{r} \partial_r \Psi) - \frac{2}{\beta} \frac{1}{r}B_0 \partial_z (\nabla^2 A - \frac{2}{r} \partial_r A) - \frac{2}{r}u_0 \partial_z u_\phi + \frac{2}{\beta} \frac{2}{r^2}B_0 \xi \partial_z B_\phi \nonumber\\
& - \frac{1}{\reye} \left[ \nabla^2 (\frac{1}{r} \nabla^2 \Psi - \frac{2}{r^2}\partial_r\Psi) - \frac{1}{r^3} \nabla^2 \Psi + \frac{2}{r^4}\partial_r\Psi\right] = N^{(\Psi)} \label{eq:Psi_perturbed}
%\end{split}
\end{align}

@@

We solve the equations at each order using Dedalus, an open source pseudospectral code. We solve the radial portion of the eigenvectors on a basis of Chebyshev polynomials subject to our radial boundary conditions. %Because our domain is bounded only radially and not vertically, we solve the radial dimension on a basis of Chebyshev polynomials, and the vertical dimension on a Fourier series basis. 
We solve Equation \ref{eq:ordere} as a linear eigenvalue problem, and Equation \ref{eq:ordere2} as a linear boundary value problem. 
 %should i mention spurious eigenvals? 
%To analyze the MRI system at marginality, we fix the parameters listed in Table \ref{table:parameters} and determine the critical $\reym$ and $k_z$ by repeatedly solving Equation \ref{eq:ordere} to determine the smallest parameter values for which the fastest growing mode is zero. That is, we solve the linear eigenvalue problem for eigenvalues $\sigma = \gamma + i \omega$. Figure \ref{fig:growth_rates} shows linear MRI growth rates $\gamma$ in the $(\reym, k_z)$ plane. For the fiducial standard MRI parameters in Table \ref{table:parameters} we find critical parameters $\reym_c = 0.84$, $k_c = 0.0127$.

%We defer full expressions and derivations thereof to the Appendix, and focus here on features of the wide gap system which distinguish it from the narrow gap limit.

The result of the weakly nonlinear analysis is a single amplitude equation for $\alpha$. This amplitude equation is found by enforcing a solvability condition on Equation \ref{eq:ordere3}. We find

@@

%For the standard MRI we derive a real GLE, and for the helical MRI we derive a complex GLE. 
 %This complex GLE is a natural form for the HMRI, as it is a traveling wave instability.

@@

%\begin{figure*}
%\centering
%\includegraphics[width=\textwidth]{../figures/fiducial_standard_mri_stressplot_Pm_16E-7_cc.png}
%\caption{Reynolds ($\mathbb{T}_{R} = u_r u_\phi$), Maxwell ($\mathbb{T}_{M} = -\frac{2}{\beta} B_r B_\phi$), %and total stress ($\mathbb{T} = \mathbb{T}_{R} + \mathbb{T}_{M}$) for the fiducial standard MRI case. }
%\end{figure*}

@@

%\begin{figure}
%\centering
%\includegraphics[width=\columnwidth]{../figures/widegap_3D_vel_screenshot.png}
%\caption{Visualization of the velocity perturbation field in the fiducial SMRI case. The three dimensional path of t%wenty randomly placed tracer particles is plotted, where the particle velocity is defined as the total (first + second %order) velocity perturbation. This is meant to be a visualization only of the velocity field, so the particle motion f%eels no contribution from the magnetic field.}
%\end{figure}

\subsection{Standard MRI}
\label{sec:SMRI}
For the standard MRI we derive a real GLE. Here we note a departure from the behavior of the narrow gap system. The purely conducting boundary condition states that the axial component of the current ($\mathbf{J}_z = [\nabla \times \mathbf{B}]_z)$ must be zero at the walls. In the thin gap geometry, the purely conducting boundary condition on the azimuthal magnetic field is $\partial_x(B_y) = 0$ for axisymmetric perturbations. A spatially constant azimuthal field satisfies both the thin-gap MRI equations and this boundary condition. This neutral mode is formally included in the analysis of \citet{Umurhan:2007hs} and yields a second amplitude equation in the form of a simple diffusion equation. This amplitude equation decouples from the GLE because of the translational symmetry of the thin-gap geometry. Because that symmetry is not preserved in the wide-gap case, \citet{Umurhan:2007hs} postulate that slow variation in the wide-gap geometry will be governed by two coupled amplitude equations. %However, the wide-gap conducting boundary condition $J_z = \frac{1}{r} \partial_r (r B_\phi) = 0$, together with the purely geometric term in Equation \ref{eq:Bphi_perturbed}, conspire to prevent the wide-gap geometry from sustaining a neutral mode. 
However, the purely geometric term in Equation \ref{eq:Bphi_perturbed} prevents the wide-gap geometry from sustaining a neutral mode. We note that a neutral mode of the form $B_\phi(r) \propto \frac{1}{r}$ would exist in a resistance-free approximation. Here, however, this mode does not exist and we derive a single real GLE as the amplitude equation of the standard MRI.

@@

The form of the nonlinear terms, detailed in Appendix \ref{app:nonlinear_terms}, represent a departure from the thin-gap theory. The narrow gap nonlinear terms at both second and third orders are linear combinations of Jacobians. The nonlinear terms in the wide-gap case differ from their thin-gap analogues with the addition of vertical advective terms. These terms derive from the advective derivatives in the momentum and induction equations, but are filtered out in the thin-gap approximation. The nonlinear terms ultimately determine the saturation amplitude of the system, as described below. %These advective terms allow the nonlinear contributions at both second and third order (i.e. N$_2$ and N$_3$) not to individually satisfy the boundary conditions on $\Psi$ and $u$. %(I'm just guessing but this may lead to higher torque on inner cylinder -- should calculate)

@@

We examine the behavior of the wide gap MRI system as a function of $\Pm$ in the regime $\Pm \ll 1$. Figure \ref{fig:coefficients} shows the critical parameters $k_c$ and $\reym$ as a function of $\Pm$, as well as the GLE linear coefficient $b$ and the diffusion coefficient $d$. From Equation \ref{eq:gle} it is readily apparent that the asymptotic saturation amplitude is $\alpha_{s} = \pm \sqrt{b/c}$, and we plot the dependence of $\alpha_{s}$ on $\Pm$ in the bottom panel of \ref{fig:coefficients}. %The GLE $b$ and $d$ coefficients are insensitive to $\Pm$, so the $\alpha_{s}$ scaling is driven by the dependence of the nonlinear coefficient $c$ on $\Pm$. 
Note that because $\reym$ is essentially constant as a function of $\Pm$, the saturation amplitude is equivalently sensitive to $\reye^{-1}$. We find by fitting the data that the saturation amplitude scales as $\alpha_{s} \sim \Pm^{0.777}$. For these same boundary conditions, \citeauthor{Umurhan:2007hs} (\citeyear{Umurhan:2007hs}) find that the narrow gap saturation amplitude scales as $\Pm^{2/3}$. They find that this amplitude dependence is driven by the $\Pm^{1/3}$ dependence of the linear boundary layer. Boundary layer analysis similarly reveals a $\nu^{1/3}$ dependence for the radial extent of the boundary layer in TC flow (\citei{Goodman:2002ix}). %Why does the $\Pm$ scaling of the boundary layer width not translate to a stronger $\Pm$ scaling for the saturation amplitude in the wide gap case? The boundary layer in the wide gap case is strongly localized at the walls, i.e. $r_1$ and $r_2$. 
Figure \ref{fig:thin_wide_comparison} shows the structure of the third-order nonlinear term $N_{31}^{(A)}$ as a function of $\Pm$ for both the narrow and wide gap standard MRI. $N_{31}$ is the vector that determines the GLE coefficient $c$, and thus the scaling of the saturation amplitude because of the insensitivity of $b$ to $\Pm$ (see Appendix \ref{app:basic_equations} for the wide gap case, and \citei{Umurhan:2007hs}, \citei{Clark:2016} for the narrow gap equations). Clearly, the width of the boundary layers scales with $\Pm$ in both the wide and narrow gap MRI. This translates to a steeper saturation amplitude $\Pm$ dependence in the wide gap case.%However, in the narrow gap case this scaling extends prominently into the center of the channel, whereas for the wide gap case the bulk of the flow is relatively unaffected by the boundary layer scaling.

@@

%\section{Channel Modes}
%Channel modes are radially independent linear MRI modes which are, under certain conditions, exact nonlinear solutions to the MRI equations.

@@

The phase dynamics of the complex GLE are well-studied in a variety of systems, and depend on the values of the GLE coefficients. The complex GLE may be unstable to traveling wave instabilities such as the Benjamin-Feir instability, a generalization of the Ekhaus instability. The complex GLE can also admit spatiotemporal chaos, and various classes of coherent structures (\citei{Aranson:2002}).
%Because the saturation amplitude of the weakly nonlinear helical MRI is governed by a complex GLE, the saturated state of the helical MRI may be unstable to 
Although a detailed description of the phase dynamics in the helical MRI is beyond the scope of this work, we note that such long-wavelength, long-timescale behavior may be observed in liquid metal helical MRI experiments.

@@

Our work should be placed in the broader context of emergent pattern formation in physical systems. The real Ginzburg-Landau equation derived here governs the slow-parameter evolution of the standard MRI close to threshold. The GLE arises in a number of other physical systems, and in each case it is a consequence not of the particular physics at hand, but of the underlying symmetries in the problem. Here we make a phenomenological comparison to two other systems that give rise to a GLE. The first and perhaps most famous is Rayleigh-B\'enard convection, in which a fluid between two plates is heated from below (\citei{Newell:1969wr}). If we take the plane of the fluid to be infinite in the horizontal plane, the system is initially translationally symmetric. %When the system eventually becomes unstable to convective heat transfer it undergoes a symmetry breaking, allowing the 
At the onset of convection the system undergoes a symmetry breaking, forming rolls, or convection cells, which break the horizontal translational invariance. Analogously, the standard MRI system considered here is initially vertically translationally symmetric, because we idealize the TC device as an infinitely long cylinder. The MRI breaks this symmetry, forming cells along the vertical length of the domain. Just as Rayleigh-B\'enard cells transport heat vertically, the MRI cells transport angular momentum horizontally. The symmetry breaking of each of these systems is described near onset by the real GLE.

@@

%As this formalism is extended to more astrophysically relevant conditions  
%The application and extension of the formalism presented here to more astrophysically representative conditions will allow an analytical understanding of weakly nonlinear MRI behavior.

%We find that the saturation amplitude of the standard MRI scales somewhat more steeply with $\Pm$ than does this same system system in the narrow gap approximation.

@@

Here we detail the perturbation analysis described in Section \ref{sec:wnl_analysis}. The linear system is described by Equation \ref{eq:unperturbed_matrix_eqns}, where

@@

%The GLE is derived from 
%Equation \ref{eq:ordere3} is resonant with 
To find a bounded solution at $\mathcal{O}(\epsilon^3)$ we must eliminate secular terms: terms which are resonant with the solution to the linear homogenous equation $(\mathcal{L} + \xi \widetilde{\mathcal{H}}) \mathbf{V} = 0$ and cause the solution to grow without bound. Secular terms in our system are those that are resonant with the linear ansatz (Equation \ref{eq:ansatz}), i.e. terms with $e^{i k_c z}$ z-dependence. To eliminate these terms we enforce a solvability condition, which arises from a corollary to the Fredholm alternative. The Fredholm alternative states that if we consider a system of equations $\mathcal{L} \mathbf{V} = \mathbf{b}$ and its adjoint homogenous system $\mathcal{L}^\dagger \mathbf{V}^\dagger = 0$, only one of two conditions holds. Either there exists one and only one solution to the inhomogenous system, or the homogenous adjoint equation has a nontrivial solution. The relevant corollary arises as a consequence of the second condition: if $\mathcal{L}^\dagger \mathbf{V}^\dagger = 0$ has a nontrivial solution, then $\mathcal{L} \mathbf{V} = \mathbf{b}$ has a solution if and only if $\langle \mathbf{V}^\dagger | b \rangle = 0$.

% This solvability condition is the requirement that the inner product of the equation with its adjoint homogenous solution is zero. 
We define the adjoint operator $\mathcal{L}^\dagger$ and solution $\mathbf{V}^\dagger$ as

@@

%\bibitem[{Chandrasekhar(1960)}]{Chandrasekhar:1960wh}
%Chandrasekhar, S. 1960, PNAS 46, 253

>>>>

\documentclass{emulateapj}
\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}

@@

\usepackage[colorlinks,urlcolor=blue,citecolor=blue,linkcolor=blue]{hyperref}
\usepackage{gensymb}
\usepackage{amsmath}
\usepackage{color}
\usepackage[titletoc,title]{appendix}

@@

\begin{document}
\title{The Weakly Nonlinear Magnetorotational Instability in a Global, Cylindrical Taylor-Couette Flow} 
\author{Susan E. Clark\altaffilmark{1}}
\author{Jeffrey S. Oishi\altaffilmark{2,}\altaffilmark{3}}
\altaffiltext{1}{Department of Astronomy, Columbia University, New York, NY} 
\altaffiltext{2}{Department of Astrophysics, American Museum of Natural History, New York, NY}
\altaffiltext{3}{Department of Physics and Astronomy, Bates College}

@@

\begin{table*}[ht]
\normalsize
\caption{Fiducial parameters for MRI runs} \label{table:parameters}
\centering
\begin{tabular}{rlrrrrrrr}
  \hline
 & $\xi$ & $\Pm$ & $\beta$ & $\Omega_2/\Omega_1$ & $R_1/R_2$ & radial magnetic b.c. \\ 
  \hline\hline
Standard MRI & 0 & 1.6E-6 & 41.2 & 0.121 & 0.33 & conducting \\ 
Helical MRI & 4 & 1E-6 & 1.7E-2 & 0.27 & 0.5 & insulating\\ 
   \hline
\end{tabular}
\end{table*}

@@

@@

\beq\label{momentum}
\partial_t \mathbf{u} + \mathbf{u} \cdot \nabla \mathbf{u} = -\frac{1}{\rho}\nabla P - \nabla\Phi + \frac{1}{c\rho} \left(\mathbf{J}\times\mathbf{B}\right) + \nu\nabla^2 \mathbf{u} 
\eeq

@@

These definitions automatically satisfy Equations \ref{eq:incompressibility} and \ref{eq:solenoid} for axisymmetric disturbances. We note that in the linearized equations, streamfunctions of the form $u_x = \partial_z \Psi$, $u_z = -(\partial_r + \frac{1}{r}) \Psi$, and the corresponding definitions of the magnetic vector potential, are convenient choices, but we define Equations \ref{eq:stokes} and \ref{eq:fluxfunc} for this nonlinear investigation because of the incommutability of $\partial_r$ and $\partial_r + \frac{1}{r}$.

@@

\begin{align}
& \frac{1}{r}\partial_t (\nabla^2 \Psi - \frac{2}{r} \partial_r \Psi) - \frac{2}{\beta} \frac{1}{r}B_0 \partial_z (\nabla^2 A - \frac{2}{r} \partial_r A) - \frac{2}{r}u_0 \partial_z u_\phi + \frac{2}{\beta} \frac{2}{r^2}B_0 \xi \partial_z B_\phi \nonumber\\
& - \frac{1}{\reye} \left[ \nabla^2 (\frac{1}{r} \nabla^2 \Psi - \frac{2}{r^2}\partial_r\Psi) - \frac{1}{r^3} \nabla^2 \Psi + \frac{2}{r^4}\partial_r\Psi\right] = N^{(\Psi)} \label{eq:Psi_perturbed}
\end{align}

@@

We solve the equations at each order using Dedalus, an open source pseudospectral code. We solve the radial portion of the eigenvectors on a basis of Chebyshev polynomials subject to our radial boundary conditions
We solve Equation \ref{eq:ordere} as a linear eigenvalue problem, and Equation \ref{eq:ordere2} as a linear boundary value problem. 
The result of the weakly nonlinear analysis is a single amplitude equation for $\alpha$. This amplitude equation is found by enforcing a solvability condition on Equation \ref{eq:ordere3}. We find

@@

@@

@@

\subsection{Standard MRI}
\label{sec:SMRI}
For the standard MRI we derive a real GLE. Here we note a departure from the behavior of the narrow gap system. The purely conducting boundary condition states that the axial component of the current ($\mathbf{J}_z = [\nabla \times \mathbf{B}]_z)$ must be zero at the walls. In the thin gap geometry, the purely conducting boundary condition on the azimuthal magnetic field is $\partial_x(B_y) = 0$ for axisymmetric perturbations. A spatially constant azimuthal field satisfies both the thin-gap MRI equations and this boundary condition. This neutral mode is formally included in the analysis of \citet{Umurhan:2007hs} and yields a second amplitude equation in the form of a simple diffusion equation. This amplitude equation decouples from the GLE because of the translational symmetry of the thin-gap geometry. Because that symmetry is not preserved in the wide-gap case, \citet{Umurhan:2007hs} postulate that slow variation in the wide-gap geometry will be governed by two coupled amplitude equations. However, the purely geometric term in Equation \ref{eq:Bphi_perturbed} prevents the wide-gap geometry from sustaining a neutral mode. We note that a neutral mode of the form $B_\phi(r) \propto \frac{1}{r}$ would exist in a resistance-free approximation. Here, however, this mode does not exist and we derive a single real GLE as the amplitude equation of the standard MRI.

@@

The form of the nonlinear terms, detailed in Appendix \ref{app:nonlinear_terms}, represent a departure from the thin-gap theory. The narrow gap nonlinear terms at both second and third orders are linear combinations of Jacobians. The nonlinear terms in the wide-gap case differ from their thin-gap analogues with the addition of vertical advective terms. These terms derive from the advective derivatives in the momentum and induction equations, but are filtered out in the thin-gap approximation. The nonlinear terms ultimately determine the saturation amplitude of the system, as described below.

@@

We examine the behavior of the wide gap MRI system as a function of $\Pm$ in the regime $\Pm \ll 1$. Figure \ref{fig:coefficients} shows the critical parameters $k_c$ and $\reym$ as a function of $\Pm$, as well as the GLE linear coefficient $b$ and the diffusion coefficient $d$. From Equation \ref{eq:gle} it is readily apparent that the asymptotic saturation amplitude is $\alpha_{s} = \pm \sqrt{b/c}$, and we plot the dependence of $\alpha_{s}$ on $\Pm$ in the bottom panel of \ref{fig:coefficients}. 
Note that because $\reym$ is essentially constant as a function of $\Pm$, the saturation amplitude is equivalently sensitive to $\reye^{-1}$. We find by fitting the data that the saturation amplitude scales as $\alpha_{s} \sim \Pm^{0.777}$. For these same boundary conditions, \citeauthor{Umurhan:2007hs} (\citeyear{Umurhan:2007hs}) find that the narrow gap saturation amplitude scales as $\Pm^{2/3}$. They find that this amplitude dependence is driven by the $\Pm^{1/3}$ dependence of the linear boundary layer. Boundary layer analysis similarly reveals a $\nu^{1/3}$ dependence for the radial extent of the boundary layer in TC flow (\citei{Goodman:2002ix}).
Figure \ref{fig:thin_wide_comparison} shows the structure of the third-order nonlinear term $N_{31}^{(A)}$ as a function of $\Pm$ for both the narrow and wide gap standard MRI. $N_{31}$ is the vector that determines the GLE coefficient $c$, and thus the scaling of the saturation amplitude because of the insensitivity of $b$ to $\Pm$ (see Appendix \ref{app:basic_equations} for the wide gap case, and \citei{Umurhan:2007hs}, \citei{Clark:2016} for the narrow gap equations). Clearly, the width of the boundary layers scales with $\Pm$ in both the wide and narrow gap MRI. This translates to a steeper saturation amplitude $\Pm$ dependence in the wide gap case.

@@

@@

The phase dynamics of the complex GLE are well-studied in a variety of systems, and depend on the values of the GLE coefficients. The complex GLE may be unstable to traveling wave instabilities such as the Benjamin-Feir instability, a generalization of the Ekhaus instability. The complex GLE can also admit spatiotemporal chaos, and various classes of coherent structures (\citei{Aranson:2002}).Although a detailed description of the phase dynamics in the helical MRI is beyond the scope of this work, we note that such long-wavelength, long-timescale behavior may be observed in liquid metal helical MRI experiments.

@@

Our work should be placed in the broader context of emergent pattern formation in physical systems. The real Ginzburg-Landau equation derived here governs the slow-parameter evolution of the standard MRI close to threshold. The GLE arises in a number of other physical systems, and in each case it is a consequence not of the particular physics at hand, but of the underlying symmetries in the problem. Here we make a phenomenological comparison to two other systems that give rise to a GLE. The first and perhaps most famous is Rayleigh-B\'enard convection, in which a fluid between two plates is heated from below (\citei{Newell:1969wr}). If we take the plane of the fluid to be infinite in the horizontal plane, the system is initially translationally symmetric. 
At the onset of convection the system undergoes a symmetry breaking, forming rolls, or convection cells, which break the horizontal translational invariance. Analogously, the standard MRI system considered here is initially vertically translationally symmetric, because we idealize the TC device as an infinitely long cylinder. The MRI breaks this symmetry, forming cells along the vertical length of the domain. Just as Rayleigh-B\'enard cells transport heat vertically, the MRI cells transport angular momentum horizontally. The symmetry breaking of each of these systems is described near onset by the real GLE.

@@

@@

Here we detail the perturbation analysis described in Section \ref{sec:wnl_analysis}. The perturbation series is described by Equations \ref{eq:ordere} - \ref{eq:ordere3}, where

@@

To find a bounded solution at $\mathcal{O}(\epsilon^3)$ we must eliminate secular terms: terms which are resonant with the solution to the linear homogenous equation $(\mathcal{L} + \xi \widetilde{\mathcal{H}}) \mathbf{V} = 0$ and cause the solution to grow without bound. Secular terms in our system are those that are resonant with the linear ansatz (Equation \ref{eq:ansatz}), i.e. terms with $e^{i k_c z}$ z-dependence. To eliminate these terms we enforce a solvability condition, which arises from a corollary to the Fredholm alternative. The Fredholm alternative states that if we consider a system of equations $\mathcal{L} \mathbf{V} = \mathbf{b}$ and its adjoint homogenous system $\mathcal{L}^\dagger \mathbf{V}^\dagger = 0$, only one of two conditions holds. Either there exists one and only one solution to the inhomogenous system, or the homogenous adjoint equation has a nontrivial solution. The relevant corollary arises as a consequence of the second condition: if $\mathcal{L}^\dagger \mathbf{V}^\dagger = 0$ has a nontrivial solution, then $\mathcal{L} \mathbf{V} = \mathbf{b}$ has a solution if and only if $\langle \mathbf{V}^\dagger | b \rangle = 0$.

We define the adjoint operator $\mathcal{L}^\dagger$ and solution $\mathbf{V}^\dagger$ as

@@
